Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-07-11 introduction.tex
ex.1 introduction

The latter case is related to a motivation  of  the present work:
automated music transcription. Representations that capture  music performances
are essentially linear:   
audio files, or the widely used MIDI file format~\cite{?}.
Such representations ignore the hierarchical structures that frame the
conception of music, at least in the western area. These structures, on the other hand,
are present, either explicitly  or implicitly,
in music notation~\cite{Gould11Notation}: music scores are partitioned in measures, measures
in beats, and beats can be further recursively divided.
It follows that music events do not occur at arbitrary timestamps,
but respect a discrete partitioning of the  timeline incurred by
these recursive divisions.
The \emph{transcription problem} takes
as input a linear representation (audio or MIDI) and aims at re-constructing
these structures
by mapping input events to this hierarchical rhythmic space.
It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990}
over an infinite alphabet of timed events.

@@

\begin{example}[Running example]\label{ex:running}
We
illustrate our framework with music transcription: given
a \emph{timeline} of musical events with timestamps from an infinite
domain as input, parse it
into a structured music score. Input events
are pairs $\< \mu, \tau>$, with $\mu \in \Sigma$, where $\Sigma$ stands for the
set of MIDI message symbols~\cite{?}
and  $\tau \in \mathbb{Q}$ is a timestamp. Such inputs typically correspond
to the recording of a live performance. The output of parsing
is a sequence of events (or \emph{notes}) $\< \nu, \tau'>$   in
Common Western Music Notation (CWMN)~\cite{Gould11Notation}
where event symbols $\nu$ belong to the domain $\Delta$
of \emph{pitches} (e.g., A4, G5, etc.), and timestamp $\tau'$
belongs to a rhythmic ``grid'' obtained from recursive divisions:
whole notes ($\musWhole$) splitted in halves ($\musHalf$), halves
in quarters ($\musQuarter$), eights ($\musEighth$), etc.
The following concrete sequences
will be used:
\begin{enumerate}
  \item A performance  $I = [ \<\mu_1, 0.07>, \<\mu_2, 0.72>, \<\mu_3, 0.91>, \<\mu_4, 1.05>, \<\mu_5, 1.36>, \<\mu_6, 1.71>]$,
     over interval $[0,2[$,
  \item A score $O =$ \includegraphics[scale=0.20]{pictures/score5.png}, corresponding to
      a hierarchical structures (measures, tuplets) that can be linearized as a sequence of events
      $[\<\nu_1,0>, \<\nu_2,\frac{3}{4}>, \<\nu_3,\frac{7}{8}>, \<\nu_4,1>, \<\nu_5,\frac{4}{3}>, \<\nu_6\frac{5}{3}>]$
\end{enumerate}
\smallskip

We will show that $O$ is a soution for the
parsing of $I$. Note that several other parsings (e.g.,
\includegraphics[scale=0.20]{pictures/score4.png}) are possible.
SW-parsing associates a weight value
to each solution, and our framework
aims at selecting the best one with respect to this weight.
\endex
\end{example}

>>>>

The latter case is related to a motivation  of  the present work:
automated music transcription. Representations that capture  music performances
are essentially linear:   
audio files, or the widely used MIDI file format~\cite{MIDIfile}.
Such representations ignore the hierarchical structures that frame the
conception of music, at least in the western area. These structures, on the other hand,
are present, either explicitly  or implicitly,
in Common Western Music Notation~\cite{Gould11Notation}
%in music notation~\cite{Gould11Notation}: 
music scores are partitioned in measures, 
measures in beats, and beats can be further recursively divided.
It follows that music events do not occur at arbitrary timestamps,
but respect a discrete partitioning of the  timeline incurred by
these recursive divisions.
The \emph{transcription problem} takes
as input a linear representation (audio or MIDI) and aims at re-constructing
these structures
by mapping input events to this hierarchical rhythmic space.
It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990}
over an infinite alphabet of timed events.

@@

\begin{example}[Running example]\label{ex:running}
We illustrate our framework with music transcription: given
a \emph{timeline} of timestamped musical events from an infinite
alphabet $\Sigma$ as input, 
parse it into a structured music score. 
Input events of $\Sigma$
are pairs $\< \mu, \tau>$, were $\mu$ is a MIDI message symbol~\cite{MIDIfile}
and $\tau \in \mathbb{Q}$ is a timestamp. Such inputs typically correspond
to the recording of a live performance. The output of parsing
is a sequence of events (or \emph{notes}) $\< \nu, \tau'>$ 
in an alphabet $\Delta$
%in Common Western Music Notation (CWMN)~\cite{Gould11Notation}
where event symbols $\nu$ are \emph{pitch names} (e.g., A4, G5, \etc.), 
and timestamp $\tau'$ belongs to a rhythmic ``grid'' obtained from recursive divisions:
whole notes ($\musWhole$) splitted in halves ($\musHalf$), halves
in quarters ($\musQuarter$), eights ($\musEighth$), etc.
The following concrete sequences
will be used:
\begin{enumerate}
  \item A performance  $I = [ \<49, 0.07>, \<51, 0.72>, \<52, 0.91>, \<54, 1.05>, \<56, 1.36>, \<57, 1.71>]$, % over interval $[0,2[$,
  \item A score $O =$ 
  \includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex1.pdf},
  %\includegraphics[scale=0.20]{pictures/score5.png}, 
  corresponding to a hierarchical structures (measures, tuplets) 
  that can be linearized as a sequence of events
      $[\<\mathrm{A}4, 0>, 
        \<\mathrm{B}4, \frac{3}{4}>, 
        \<\mathrm{C}5, \frac{7}{8}>, 
        \<\mathrm{D}5, 1>, 
        \<\mathrm{E}5,\frac{4}{3}>, 
        \<\mathrm{F}5\frac{5}{3}>]$
\end{enumerate}
\smallskip

We will show that $O$ is a soution for the
parsing of $I$. Note that several other parsings (e.g.,
\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}) are possible.
SW-parsing associates a weight value
to each solution, and our framework
aims at selecting the best one with respect to this weight.
\endex
\end{example}
